---
title: "Appendix A - Geographe Marine Park Natural Values Report"
subtitle: "South-west Marine Parks Network"
filters:
  - figure-numbering.lua
format:
  pdf:
    output-file: "Project 4.21-Geographe-2-Appendix A-q-Natural values.pdf"
    fig-pos: 'H'
    geometry: 
      - top=2cm
      - bottom=2.5cm
      - left=2cm
      - right=2cm
    include-in-header:
      text: |
        \usepackage{caption}
        \usepackage{float}
        \usepackage{chngcntr}
        \newcounter{metric}
        \counterwithin*{figure}{metric}
        \captionsetup{format=plain, indention=0pt}
  html:
    output-file: "Project 4.21-Geographe-2-Appendix A-q-Natural values.html"
    fig-dpi: 96
toc: true
execute:
  echo: false
  warning: false
  message: false
post-render: 10_move-pdf.R
---

```{r}
#| include: false
library(here)
library(ggplot2)
library(knitr)
library(terra)
library(tidyterra)
library(cowplot)
library(png)
library(grid)
library(tidytext)
library(ggtext)
```

# Natural Values

Understanding of natural values

## Benthic ecosystem extent
### Observations

```{r}
#| fig-width: 8
#| fig-height: 4
#| fig-cap: "Habitat classes observed in stereo-BRUV imagery. Habitat classes observed in spatially balanced stereo-BRUV and drop-camera imagery visualised as spatial pie charts. State and Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
p_scatterpie_multi <- readRDS(here("plots/geographe/habitat/GeographeAMP_scatterpie_2014-2024.rds"))
p_scatterpie_multi
```

{{< pagebreak >}}

### Dominant habitat types

```{r}
#| fig-width: 8
#| fig-height: 5
#| fig-cap: "Predicted dominant habitat type from stereo-BRUV groundtruthing.The green polygon represents the extent of no-take National Park Zones and State no-take sanctuary zones. State and Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
p_cat_multi <- readRDS(here("plots/geographe/habitat/GeographeAMP_predicted-habitat-categorical_2014-2024.rds"))
p_cat_multi
```

{{< pagebreak >}}

### Habitat distribution and probability

```{r}
#| fig-width: 8
#| fig-height: 7
#| fig-cap: "Predicted habitat distribution probability and uncertainty from stereo-BRUV groundtruthing. The green polygon represents the extent of no-take National Park Zones. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
img <- png::readPNG(here("plots/geographe/habitat/GeographeAMP_predicted-dominant-benthos-and-combined-se_2014-2024.png"))
grid::grid.raster(img)
```

{{< pagebreak >}}

### Funcional reef distribution and probability

```{r}
#| fig-width: 8
#| fig-height: 7
#| fig-cap: "Predicted reef distribution probability and uncertainty from stereo-BRUV groundtruthing.The green polygon represents the extent of no-take National Park Zones. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
img <- png::readPNG(here("plots/geographe/habitat/GeographeAMP_predicted-reef-and-se_2014-2024.png"))
grid::grid.raster(img)
```

{{< pagebreak >}}

## Benchmarks of benthic natural values

```{=latex}
% Continue the running figure number into the benchmark sections, then switch
% to <metric>.<plot> numbering. `metric` is seeded to the last benthic-extent
% figure number so the majors carry on sequentially (4, 5, 6, ...).
\setcounter{metric}{\value{figure}}
\renewcommand{\thefigure}{\arabic{metric}.\arabic{figure}}
```

### Seagrass

```{=latex}
\stepcounter{metric}
```


```{r}
#| fig-width: 6
#| fig-height: 2.4
#| fig-cap: "Seagrass: temporal plots by depth contour from stereo-BRUV groundtruthing. Dashed vertical line indicates the date of establishment of the marine park. GMP = GMP (Commonwealth), HPZ = Habitat Protection Zone, NPZ = National Park Zone, NCMP = Ngari Capes Marine Park (State), SZ = Sanctuary Zone."
p_control <- readRDS(here("plots/geographe/habitat/GeographeAMP_control-plot_seagrass.rds"))
p_control
```



```{r}
#| fig-width: 8
#| fig-height: 4.7
#| fig-cap: "Seagrass: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. State and Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
img <- png::readPNG(here("plots/geographe/habitat/GeographeAMP_predicted-individual-habitat_seagrass_2014-2024.png"))
grid::grid.raster(img)
```

{{< pagebreak >}}

### Macroalgae

```{=latex}
\stepcounter{metric}
```


```{r}
#| fig-width: 6
#| fig-height: 2.4
#| fig-cap: "Macroalgae: temporal plots by depth contour from stereo-BRUV groundtruthing. Dashed vertical line indicates the date of establishment of the marine park. GMP = GMP (Commonwealth), HPZ = Habitat Protection Zone, NPZ = National Park Zone, NCMP = Ngari Capes Marine Park (State), SZ = Sanctuary Zone."
p_control <- readRDS(here("plots/geographe/habitat/GeographeAMP_control-plot_macroalgae.rds"))
p_control
```


```{r}
#| fig-width: 8
#| fig-height: 4.7
#| fig-cap: "Macroalgae: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. State and Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
img <- png::readPNG(here("plots/geographe/habitat/GeographeAMP_predicted-individual-habitat_macroalgae_2014-2024.png"))
grid::grid.raster(img)
```

{{< pagebreak >}}

### Sessile invertebrates

```{=latex}
\stepcounter{metric}
```


```{r}
#| fig-width: 6
#| fig-height: 2.4
#| fig-cap: "Sessile invertebrates: temporal plots by depth contour from stereo-BRUV groundtruthing. Dashed vertical line indicates the date of establishment of the marine park. GMP = GMP (Commonwealth), HPZ = Habitat Protection Zone, NPZ = National Park Zone, NCMP = Ngari Capes Marine Park (State), SZ = Sanctuary Zone."
p_control <- readRDS(here("plots/geographe/habitat/GeographeAMP_control-plot_sessile-invertebrates.rds"))
p_control
```


```{r}
#| fig-width: 8
#| fig-height: 4.7
#| fig-cap: "Sessile invertebrates: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. State and Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
img <- png::readPNG(here("plots/geographe/habitat/GeographeAMP_predicted-individual-habitat_sessile-invertebrates_2014-2024.png"))
grid::grid.raster(img)
```

## Benchmarks of demersal fish natural values

### Survey effort

```{=latex}
\stepcounter{metric}
\renewcommand{\thefigure}{\arabic{metric}}
```

```{r}
#| fig-width: 8
#| fig-height: 8
#| fig-cap: "Spatial distribution of observed species richness and total abundance (MaxN) per stereo-BRUV drop, pooled across all deployments. State and Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters."
p_bubble <- readRDS(here("plots/geographe/fish/GeographeAMP_bubbleplot_richness-abundance.rds"))
p_bubble
```

{{< pagebreak >}}

```{=latex}
\renewcommand{\thefigure}{\arabic{metric}.\arabic{figure}}
```

### Species richness

```{=latex}
\stepcounter{metric}
```

```{r}
#| fig-width: 6
#| fig-height: 3
#| fig-cap: "Demersal fish - Species richness: benchmark plots by zone and depth contour from stereo-BRUV groundtruthing. Dashed vertical line indicates the date of establishment of the marine park. GMP = GMP (Commonwealth), HPZ = Habitat Protection Zone, NPZ = National Park Zone, NCMP = Ngari Capes Marine Park (State), SZ = Sanctuary Zone."
p_control <- readRDS(here("plots/geographe/fish/GeographeAMP_control-plot_species-richness-per-bruv.rds"))
p_control
```

```{r}
#| fig-width: 8
#| fig-height: 3
#| fig-cap: "Demersal fish - Species richness: species accumulation curve by zone type from stereo-BRUV groundtruthing."
p_bar <- readRDS(here("plots/geographe/fish/GeographeAMP_SAC-sample.rds"))
p_bar
```

{{< pagebreak >}}

:::{.content-hidden when-format="html"}
### Species richness {.unnumbered}
:::

```{r}
#| fig-width: 8
#| fig-height: 6
#| fig-cap: "Demersal fish - Species richness: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
img <- png::readPNG(here("plots/geographe/fish/GeographeAMP_predicted-individual-fish-metric_whole-assemblage_2014-2024.png"))
grid::grid.raster(img)
```

{{< pagebreak >}}

### Total abundance

```{=latex}
\stepcounter{metric}
```



```{r}
#| fig-width: 6
#| fig-height: 3
#| fig-cap: "Demersal fish - Total abundance: benchmark plots by zone and depth contour from stereo-BRUV groundtruthing. Dashed vertical line indicates the date of establishment of the marine park. GMP = GMP (Commonwealth), HPZ = Habitat Protection Zone, NPZ = National Park Zone, NCMP = Ngari Capes Marine Park (State), SZ = Sanctuary Zone."
p_control <- readRDS(here("plots/geographe/fish/GeographeAMP_control-plot_total-abundance-per-bruv.rds"))
p_control
```

```{r}
#| fig-width: 8
#| fig-height: 3
#| fig-cap: "Demersal fish - Total abundance: top 10 most abundant species (pooled across deployments) from stereo-BRUV groundtruthing. **Bold** species are unique within the 10 most abundant species by year."
p_bar <- readRDS(here("plots/geographe/fish/GeographeAMP_top_maxn_bar_plot.rds"))
p_bar
```

{{< pagebreak >}}

:::{.content-hidden when-format="html"}
### Total abundance {.unnumbered}
:::

```{r}
#| fig-width: 8
#| fig-height: 6
#| fig-cap: "Demersal fish - Total abundance: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
img <- png::readPNG(here("plots/geographe/fish/GeographeAMP_predicted-individual-fish-metric_total-abundance_2014-2024.png"))
grid::grid.raster(img)
```

{{< pagebreak >}}

\newgeometry{top=2cm, bottom=2.1cm, left=2cm, right=2cm}

### Large Reef Fish Index\*

```{=latex}
\stepcounter{metric}
```



```{r}
#| fig-width: 6
#| fig-height: 3
#| fig-cap: "Demersal fish - Large Reef Fish Index\\*: benchmark plots by zone and depth contour from stereo-BRUV groundtruthing. Dashed vertical line indicates the date of establishment of the marine park. GMP = GMP (Commonwealth), HPZ = Habitat Protection Zone, NPZ = National Park Zone, NCMP = Ngari Capes Marine Park (State), SZ = Sanctuary Zone. \\*Large Reef Fish Index representing the biomass of all fish greater than 20 cm (B20) has been adapted here for stereo-BRUV data by removing sharks and rays."
p_control <- readRDS(here("plots/geographe/fish/GeographeAMP_control-plot_large-reef-fish-index-biomass-g-per-bruv.rds"))
p_control
```

```{r}
#| fig-width: 8
#| fig-height: 3
#| fig-cap: "Demersal fish - Large Reef Fish Index\\*: top 10 greatest biomass species by status (pooled across deployments) from stereo-BRUV groundtruthing. **Bold** species are unique within the 10 greatest biomass species by year."
p_bar <- readRDS(here("plots/geographe/fish/GeographeAMP_top_b20_bar_plot_mixed.rds"))
p_bar
```

{{< pagebreak >}}

:::{.content-hidden when-format="html"}
### Large Reef Fish Index\* {.unnumbered}
:::

```{r}
#| fig-width: 8
#| fig-height: 6
#| fig-cap: "Demersal fish - Large Reef Fish Index\\*: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
img <- png::readPNG(here("plots/geographe/fish/GeographeAMP_predicted-individual-fish-metric_large-reef-fish-index_2014-2024.png"))
grid::grid.raster(img)
```

\restoregeometry

### Reef fish thermal index

```{=latex}
\stepcounter{metric}
```



```{r}
#| fig-width: 6
#| fig-height: 3
#| fig-cap: "Demersal fish - Reef Fish Thermal Index: benchmark plots by zone and depth contour, with monthly sea surface temperature (black line, grey band = SD) overlaid, from stereo-BRUV groundtruthing. Dashed vertical line indicates the date of establishment of the marine park. GMP = GMP (Commonwealth), HPZ = Habitat Protection Zone, NPZ = National Park Zone, NCMP = Ngari Capes Marine Park (State), SZ = Sanctuary Zone."
p_control <- readRDS(here("plots/geographe/fish/GeographeAMP_control-plot_community-thermal-index-c.rds"))
p_control
```

```{r}
#| fig-width: 8
#| fig-height: 3
#| fig-cap: "Demersal fish - Reef Fish Thermal Index: species thermal niche of top 10 most abundant species (pooled across deployments) from stereo-BRUV groundtruthing. **Bold** species are unique within the 10 most abundant species by year."
p_bar <- readRDS(here("plots/geographe/fish/GeographeAMP_top_maxn_cti_bar_plot.rds"))
p_bar
```

{{< pagebreak >}}

:::{.content-hidden when-format="html"}
### Reef fish thermal index {.unnumbered}
:::

```{r}
#| fig-width: 8
#| fig-height: 6
#| fig-cap: "Demersal fish - Reef Fish Thermal Index: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
img <- png::readPNG(here("plots/geographe/fish/GeographeAMP_predicted-individual-fish-metric_cti_2014-2024.png"))
grid::grid.raster(img)
```
