---
title: "Appendix A1 - South-west Corner Marine Park (Western arm) Natural Values Report"
subtitle: "South-west Marine Parks Network"
filters:
  - figure-numbering.lua
format:
  pdf:
    output-file: "Project 4.21-SWC-western-arm-2-Appendix A1-q-Natural values.pdf"
    fig-pos: 'H'
    geometry: 
      - top=2cm
      - bottom=2.5cm
      - left=2cm
      - right=2cm
    include-in-header:
      text: |
        \usepackage{caption}
        \usepackage{float}
        \usepackage{chngcntr}
        \floatplacement{table}{H}
        \newcounter{metric}
        \counterwithin*{figure}{metric}
        \captionsetup{format=plain, indention=0pt}
        % Appendix prefix. Change A1 here and in figure-numbering.lua together.
        \renewcommand{\thefigure}{A1.\arabic{figure}}
        \renewcommand{\thetable}{A1.\arabic{table}}
  html:
    output-file: "Project 4.21-SWC-western-arm-2-Appendix A1-q-Natural values.html"
    fig-dpi: 96
    include-in-header:
      text: |
        <style>
          caption,
          .gt_caption,
          .table-caption,
          figcaption,
          .figure-caption,
          .quarto-float-caption,
          .quarto-float-caption-top,
          .quarto-float-tbl figcaption {
            text-align: justify !important;
            text-justify: inter-word;
            hyphens: auto;
          }
        </style>
    include-after-body:
      text: |
        <script>
        // Figures are prefixed by figure-numbering.lua. Quarto builds the table
        // caption after that filter runs, so the table prefix is applied here.
        // TODO Keep "A1" in step with fig_prefix in figure-numbering.lua.
        document.querySelectorAll("caption, .gt_caption, .table-caption, figcaption")
          .forEach(function (el) {
            // el.innerHTML re-serialises a non-breaking space as the literal
            // "&nbsp;" entity text, not a whitespace character, so \s alone
            // does not match it - match either form here.
            el.innerHTML = el.innerHTML.replace(/^(\s*Table(?:\s|&nbsp;)+)(\d+)/, "$1A1.$2");
          });
        </script>
toc: true
execute:
  echo: false
  warning: false
  message: false
post-render: 10_move-pdf.R
---

```{r}
#| include: false
library(here)
library(ggplot2)
library(knitr)
library(terra)
library(tidyterra)
library(cowplot)
library(png)
library(grid)
library(tidytext)
library(ggtext)
library(dplyr)
library(gt)

# Park folder and file prefix - must match 00_config.yml.
park <- "swc_westernarm"
name <- "swc_westernarm"
```

# Natural Values

Understanding of natural values

{{< pagebreak >}}

## Sampling summary

```{r}
#| label: tbl-sampling-effort
#| tbl-cap: "Summary of monitoring campaigns in the study area, showing the survey dates, sampling method, management zones sampled, and the number of samples with each data type. NPZ = National Park Zone, HPZ = Habitat Protection Zone, MUZ = Multiple Use Zone, SPZ = Special Purpose Zone."

# LaTeX will not break a line at an underscore, so a long campaign ID would run
# past its column. Swapping underscores for spaces lets the name wrap.
sampling_summary <- readRDS(here("data", park, "tidy",
                                 paste0(name, "_sampling-summary.RDS"))) %>%
  mutate(campaignid = gsub("_", " ", campaignid))

col_labels <- c("Date", "Campaign name", "Method", "Areas", "Samples",
                "Count", "Count", "Length")

sampling_summary %>%
  gt() %>%
  cols_label(dates         = col_labels[1],
             campaignid    = col_labels[2],
             method        = col_labels[3],
             areas         = col_labels[4],
             n_samples     = col_labels[5],
             habitat_count = col_labels[6],
             fish_count    = col_labels[7],
             fish_length   = col_labels[8]) %>%
  tab_spanner(label = "Habitat", columns = habitat_count) %>%
  tab_spanner(label = "Fish",    columns = c(fish_count, fish_length)) %>%
  sub_zero(zero_text = "-") %>%
  cols_width(dates         ~ pct(14),
             campaignid    ~ pct(20),
             method        ~ pct(11),
             areas         ~ pct(15),
             n_samples     ~ pct(10),
             habitat_count ~ pct(10),
             fish_count    ~ pct(10),
             fish_length   ~ pct(10)) %>%
  cols_align(align = "left",
             columns = c(dates, campaignid, areas)) %>%
  cols_align(align = "center",
             columns = c(method, n_samples, habitat_count, fish_count,
                         fish_length)) %>%
  opt_row_striping() %>%
  tab_style(style = cell_text(weight = "bold"),
            locations = list(cells_column_labels(),
                             cells_column_spanners(spanners = everything()))) %>%
  tab_options(table.width               = pct(100),
              table.font.size           = px(11),
              data_row.padding          = px(4),
              latex.use_longtable       = TRUE,
              column_labels.border.top.style = "none",
              table.border.top.style    = "none",
              table_body.hlines.style   = "none")
```

{{< pagebreak >}}

## Benthic ecosystem extent
### Observations

```{r}
#| fig-width: 8
#| fig-height: 8
#| fig-cap: "Habitat classes observed in spatially balanced stereo-BRUV and drop-camera imagery visualised as spatial pie charts. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
img <- png::readPNG(here("plots/swc_westernarm/habitat/swc_westernarm_scatterpie_2020_2021_2023_2024_2025.png"))
grid::grid.raster(img)
```

{{< pagebreak >}}

### Dominant habitat types

```{r}
#| fig-width: 8
#| fig-height: 8
#| fig-cap: "Predicted dominant habitat type from stereo-BRUV groundtruthing. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
p_cat <- readRDS(here("plots/swc_westernarm/habitat/swc_westernarm_predicted-habitat-categorical_2020_2021_2023_2024_2025.rds"))
p_cat
```

{{< pagebreak >}}

### Habitat distribution and probability

```{r}
#| fig-width: 8
#| fig-height: 7
#| fig-cap: "Predicted habitat distribution probability and uncertainty from stereo-BRUV groundtruthing. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
p_dom <- readRDS(here("plots/swc_westernarm/habitat/swc_westernarm_predicted-dominant-habitat_2020_2021_2023_2024_2025.rds"))
p_dom
```

{{< pagebreak >}}

### Funcional reef distribution and probability

```{r}
#| fig-width: 8
#| fig-height: 7
#| fig-cap: "Predicted reef distribution probability and uncertainty from stereo-BRUV groundtruthing. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
img_path <- here("plots/swc_westernarm/habitat/swc_westernarm_predicted-reef-and-se_2020-2021-2023-2024-2025.png")
img <- png::readPNG(img_path)
grid::grid.raster(img)
```

{{< pagebreak >}}

## Benchmarks of benthic natural values

```{=latex}
% Continue the running figure number into the benchmark sections, then switch
% to <metric>.<plot> numbering. `metric` is seeded to the last benthic-extent
% figure number so the majors carry on sequentially (4, 5, 6, ...).
% A1 must match fig_prefix in figure-numbering.lua.
\setcounter{metric}{\value{figure}}
\renewcommand{\thefigure}{A1.\arabic{metric}.\arabic{figure}}
```

### Sand

```{=latex}
\stepcounter{metric}
```



```{r}
#| fig-width: 8
#| fig-height: 5.5
#| fig-cap: "Sand: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. SWC = South-west Corner Marine Park (Commonwealth). Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
img <- png::readPNG(here("plots/swc_westernarm/habitat/swc_westernarm_predicted-individual-habitat_sand_2020_2021_2023_2024_2025.png"))
grid::grid.raster(img)
```

{{< pagebreak >}}

### Macroalgae

```{=latex}
\stepcounter{metric}
```



```{r}
#| fig-width: 8
#| fig-height: 5.5
#| fig-cap: "Macroalgae: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. SWC = South-west Corner Marine Park (Commonwealth). Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
img <- png::readPNG(here("plots/swc_westernarm/habitat/swc_westernarm_predicted-individual-habitat_macroalgae_2020_2021_2023_2024_2025.png"))
grid::grid.raster(img)
```

{{< pagebreak >}}

### Seagrass

```{=latex}
\stepcounter{metric}
```



```{r}
#| fig-width: 8
#| fig-height: 5.5
#| fig-cap: "Seagrass: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. SWC = South-west Corner Marine Park (Commonwealth). Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
img <- png::readPNG(here("plots/swc_westernarm/habitat/swc_westernarm_predicted-individual-habitat_seagrass_2020_2021_2023_2024_2025.png"))
grid::grid.raster(img)
```

{{< pagebreak >}}

### Sessile invertebrates

```{=latex}
\stepcounter{metric}
```



```{r}
#| fig-width: 8
#| fig-height: 5.5
#| fig-cap: "Sessile invertebrates: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. SWC = South-west Corner Marine Park (Commonwealth). Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
img <- png::readPNG(here("plots/swc_westernarm/habitat/swc_westernarm_predicted-individual-habitat_sessile-invertebrates_2020_2021_2023_2024_2025.png"))
grid::grid.raster(img)
```

{{< pagebreak >}}

### Rock

```{=latex}
\stepcounter{metric}
```



```{r}
#| fig-width: 8
#| fig-height: 5.5
#| fig-cap: "Rock: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. SWC = South-west Corner Marine Park (Commonwealth). Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
img <- png::readPNG(here("plots/swc_westernarm/habitat/swc_westernarm_predicted-individual-habitat_rock_2020_2021_2023_2024_2025.png"))
grid::grid.raster(img)
```

{{< pagebreak >}}

## Benchmarks of demersal fish natural values

### Survey effort

```{=latex}
\stepcounter{metric}
\renewcommand{\thefigure}{A1.\arabic{metric}}
```

```{r}
#| fig-width: 8
#| fig-height: 8
#| fig-cap: "Spatial distribution of observed species richness and total abundance (MaxN) per stereo-BRUV drop, pooled across all deployments. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters."
p_bubble <- readRDS(here("plots/swc_westernarm/fish/swc_westernarm_bubbleplot_richness-abundance.rds"))
p_bubble
```

{{< pagebreak >}}

```{=latex}
\renewcommand{\thefigure}{A1.\arabic{metric}.\arabic{figure}}
```

### Species richness

```{=latex}
\stepcounter{metric}
```



```{r}
#| fig-width: 6
#| fig-height: 3
#| fig-cap: "Demersal fish - Species richness: benchmark plots by zone and depth contour from stereo-BRUV groundtruthing. SWC = South-west Corner Marine Park (Commonwealth)."
p_control <- readRDS(here("plots/swc_westernarm/fish/swc_westernarm_control-plot_species-richness-per-bruv.rds"))
p_control
```

```{r}
#| fig-width: 6
#| fig-height: 3
#| fig-cap: "Demersal fish - Species richness: species accumulation curve by year from stereo-BRUV groundtruthing. SWC = South-west Corner Marine Park (Commonwealth)."
p_bar <- readRDS(here("plots/swc_westernarm/fish/swc_westernarm_SAC-sample.rds"))
p_bar
```

{{< pagebreak >}}

:::{.content-hidden when-format="html"}
### Species richness {.unlisted .unnumbered}
:::

```{r}
#| fig-width: 7
#| fig-height: 4
#| fig-cap: "Demersal fish - Species richness: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
img <- png::readPNG(here("plots/swc_westernarm/fish/swc_westernarm_predicted-individual-fish-metric_whole-assemblage_2020-2024-2025.png"))
grid::grid.raster(img)
```

{{< pagebreak >}}

### Total abundance

```{=latex}
\stepcounter{metric}
```



```{r}
#| fig-width: 6
#| fig-height: 3
#| fig-cap: "Demersal fish - Total abundance: benchmark plots by zone and depth contour from stereo-BRUV groundtruthing. SWC = South-west Corner Marine Park (Commonwealth)."
p_control <- readRDS(here("plots/swc_westernarm/fish/swc_westernarm_control-plot_total-abundance-per-bruv.rds"))
p_control
```

```{r}
#| fig-width: 8
#| fig-height: 4.5
#| fig-cap: "Demersal fish - Total abundance: top 10 most abundant species (pooled across deployments) from stereo-BRUV groundtruthing. SWC = South-west Corner Marine Park (Commonwealth)."
p_bar <- readRDS(here("plots/swc_westernarm/fish/swc_westernarm_top_maxn_bar_plot.rds"))
p_bar
```

{{< pagebreak >}}

:::{.content-hidden when-format="html"}
### Total abundance {.unlisted .unnumbered}
:::

```{r}
#| fig-width: 7
#| fig-height: 4.5
#| fig-cap: "Demersal fish - Total abundance: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
img <- png::readPNG(here("plots/swc_westernarm/fish/swc_westernarm_predicted-individual-fish-metric_total-abundance_2020-2024-2025.png"))
grid::grid.raster(img)
```

{{< pagebreak >}}

### Large Reef Fish Index\*

```{=latex}
\stepcounter{metric}
```



```{r}
#| fig-width: 6
#| fig-height: 3
#| fig-cap: "Demersal fish - Large Reef Fish Index\\*: benchmark plots by zone and depth contour from stereo-BRUV groundtruthing. SWC = South-west Corner Marine Park (Commonwealth). \\*Large Reef Fish Index representing the biomass of all fish greater than 20 cm (B20) has been adapted here for stereo-BRUV data by removing sharks and rays."
p_control <- readRDS(here("plots/swc_westernarm/fish/swc_westernarm_control-plot_large-reef-fish-index-biomass-g-per-bruv.rds"))
p_control
```

```{r}
#| fig-width: 8
#| fig-height: 4.5
#| fig-cap: "Demersal fish - Large Reef Fish Index\\*: top 10 greatest biomass species (pooled across deployments) from stereo-BRUV groundtruthing. SWC = South-west Corner Marine Park (Commonwealth)."
p_bar <- readRDS(here("plots/swc_westernarm/fish/swc_westernarm_top_b20_bar_plot_mixed.rds"))
p_bar
```

{{< pagebreak >}}

:::{.content-hidden when-format="html"}
### Large Reef Fish Index\* {.unlisted .unnumbered}
:::

```{r}
#| fig-width: 7
#| fig-height: 4
#| fig-cap: "Demersal fish - Large Reef Fish Index\\*: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
img <- png::readPNG(here("plots/swc_westernarm/fish/swc_westernarm_predicted-individual-fish-metric_large-reef-fish-index_2020-2024-2025.png"))
grid::grid.raster(img)
```

{{< pagebreak >}}

### Reef fish thermal index

```{=latex}
\stepcounter{metric}
```



```{r}
#| fig-width: 6
#| fig-height: 3
#| fig-cap: "Demersal fish - Reef Fish Thermal Index: benchmark plots by zone and depth contour, with monthly sea surface temperature (black line, grey band = SD) overlaid, from stereo-BRUV groundtruthing. SWC = South-west Corner Marine Park (Commonwealth)."
p_control <- readRDS(here("plots/swc_westernarm/fish/swc_westernarm_control-plot_community-thermal-index-c.rds"))
p_control
```

```{r}
#| fig-width: 8.5
#| fig-height: 4
#| fig-cap: "Demersal fish - Reef Fish Thermal Index: species thermal niche of top 10 most abundant species (pooled across deployments) from stereo-BRUV groundtruthing. SWC = South-west Corner Marine Park (Commonwealth)."
p_bar <- readRDS(here("plots/swc_westernarm/fish/swc_westernarm_top_maxn_cti_bar_plot.rds"))
p_bar
```

{{< pagebreak >}}

:::{.content-hidden when-format="html"}
### Reef fish thermal index {.unlisted .unnumbered}
:::

```{r}
#| fig-width: 7
#| fig-height: 5
#| fig-cap: "Demersal fish - Reef Fish Thermal Index: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
img <- png::readPNG(here("plots/swc_westernarm/fish/swc_westernarm_predicted-individual-fish-metric_cti_2020-2024-2025.png"))
grid::grid.raster(img)
```

{{< pagebreak >}}

## Threatened species

```{=latex}
\stepcounter{metric}
```

```{r}
#| label: tbl-threatened-species
#| tbl-cap: "Threatened species recorded in stereo-BRUV imagery across all survey years, showing IUCN Red List category, the number of samples each species was recorded in, and the total abundance (MaxN) summed across those samples. An asterisk denotes a species also listed as threatened under the EPBC Act 1999."

threatened_species <- readRDS(here("data", park, "tidy",
                                   paste0(name, "_threatened-species.RDS")))

threatened_species %>%
  dplyr::select(scientific_name, australian_common_name, family, listing,
                n_samples, total_maxn, years_seen) %>%
  gt() %>%
  cols_label(scientific_name        = "Species",
             australian_common_name = "Common name",
             family                 = "Family",
             listing                = "IUCN category",
             n_samples              = "Samples",
             total_maxn             = "Total MaxN",
             years_seen             = "Years recorded") %>%
  sub_missing(missing_text = "-") %>%
  cols_width(scientific_name        ~ pct(21),
             australian_common_name ~ pct(19),
             family                 ~ pct(16),
             listing                ~ pct(14),
             n_samples              ~ pct(9),
             total_maxn             ~ pct(10),
             years_seen             ~ pct(11)) %>%
  cols_align(align = "left",
             columns = c(scientific_name, australian_common_name, family)) %>%
  cols_align(align = "center",
             columns = c(listing, n_samples, total_maxn, years_seen)) %>%
  tab_style(style = cell_text(style = "italic"),
            locations = cells_body(columns = scientific_name)) %>%
  opt_row_striping() %>%
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_column_labels()) %>%
  tab_options(table.width               = pct(100),
              table.font.size           = px(11),
              data_row.padding          = px(4),
              latex.use_longtable       = TRUE,
              column_labels.border.top.style = "none",
              table.border.top.style    = "none",
              table_body.hlines.style   = "none")
```

{{< pagebreak >}}

:::{.content-hidden when-format="html"}
### Threatened species {.unlisted .unnumbered}
:::

```{r}
#| fig-width: 8
#| fig-height: 8
#| fig-cap: "Occurrence of threatened sharks and rays in stereo-BRUV imagery. Points are scaled by abundance (MaxN) and coloured by survey year; grey points show every stereo-BRUV deployment, so a panel with no coloured points is a genuine absence rather than an absence of sampling. Commonwealth and state marine park boundaries are shown. The red line delimits state and commonwealth waters."
p_threatened <- readRDS(here("plots/swc_westernarm/fish/swc_westernarm_threatened-species-occurrence_2020-2021-2023-2024-2025.rds"))
p_threatened
```

{{< pagebreak >}}

```{r}
#| fig-width: 9
#| fig-height: 10
#| fig-cap: "Occurrence of threatened sharks and rays by species (rows) and survey year (columns), page 1. Points are scaled by abundance (MaxN); grey points show every stereo-BRUV deployment, so a blank panel is a genuine absence rather than an absence of sampling."
p_threatened_grid_1 <- readRDS(here("plots/swc_westernarm/fish/swc_westernarm_threatened-species-grid_page-1.rds"))
p_threatened_grid_1
```

{{< pagebreak >}}

```{r}
#| fig-width: 9
#| fig-height: 10
#| fig-cap: "Occurrence of threatened sharks and rays by species (rows) and survey year (columns), page 2. Points are scaled by abundance (MaxN); grey points show every stereo-BRUV deployment, so a blank panel is a genuine absence rather than an absence of sampling."
p_threatened_grid_2 <- readRDS(here("plots/swc_westernarm/fish/swc_westernarm_threatened-species-grid_page-2.rds"))
p_threatened_grid_2
```

{{< pagebreak >}}

```{r}
#| fig-width: 9
#| fig-height: 10
#| fig-cap: "Occurrence of threatened sharks and rays by species (rows) and survey year (columns), page 3. Points are scaled by abundance (MaxN); grey points show every stereo-BRUV deployment, so a blank panel is a genuine absence rather than an absence of sampling."
p_threatened_grid_3 <- readRDS(here("plots/swc_westernarm/fish/swc_westernarm_threatened-species-grid_page-3.rds"))
p_threatened_grid_3
```
