---
title: "Appendix A1 - Ningaloo Marine Park Natural Values Report"
subtitle: "North-west Marine Parks Network"
filters:
  - figure-numbering.lua
format:
  pdf:
    output-file: "Project 4.21-Ningaloo-2-Appendix A1-q-Natural values.pdf"
    fig-pos: 'H'
    geometry:
      - top=2cm
      - bottom=2.5cm
      - left=2cm
      - right=2cm
    include-in-header:
      text: |
        \usepackage{caption}
        \usepackage{float}
        \usepackage{chngcntr}
        \floatplacement{table}{H}
        \newcounter{metric}
        \counterwithin*{figure}{metric}
        \captionsetup{format=plain, indention=0pt}
        % Appendix prefix. Change A here and in figure-numbering.lua together.
        \renewcommand{\thefigure}{A1.\arabic{figure}}
        \renewcommand{\thetable}{A1.\arabic{table}}
  html:
    output-file: "Project 4.21-Ningaloo-2-Appendix A1-q-Natural values.html"
    fig-dpi: 96
    include-in-header:
      text: |
        <style>
          caption,
          .gt_caption,
          .table-caption,
          figcaption,
          .figure-caption,
          .quarto-float-caption,
          .quarto-float-caption-top,
          .quarto-float-tbl figcaption {
            text-align: justify !important;
            text-justify: inter-word;
            hyphens: auto;
          }
        </style>
    include-after-body:
      text: |
        <script>
        // Figures are prefixed by figure-numbering.lua. Quarto builds the table
        // caption after that filter runs, so the table prefix is applied here.
        // TODO Keep "A" in step with fig_prefix in figure-numbering.lua.
        document.querySelectorAll("caption, .gt_caption, .table-caption, figcaption")
          .forEach(function (el) {
            // el.innerHTML re-serialises a non-breaking space as the literal
            // "&nbsp;" entity text, not a whitespace character, so \s alone
            // does not match it - match either form here.
            el.innerHTML = el.innerHTML.replace(/^(\s*Table(?:\s|&nbsp;)+)(\d+)/, "$1A1.$2");
          });
        </script>
toc: true
execute:
  echo: false
  warning: false
  message: false
post-render: 11_move-pdf.R
---

```{r}
#| include: false
library(here)
library(ggplot2)
library(knitr)
library(terra)
library(tidyterra)
library(cowplot)
library(png)
library(grid)
library(tidytext)
library(ggtext)
library(dplyr)
library(gt)

# Park folder and file prefix - must match 00_config.yml.
park <- "ningaloo"
name <- "ningaloo"
```

# Natural Values

Understanding of natural values

{{< pagebreak >}}

## Sampling summary

```{r}
#| label: tbl-sampling-effort
#| tbl-cap: "Summary of monitoring campaigns in the study area, showing the survey dates, sampling method, management zones sampled, and the number of samples with each data type. NPZ = National Park Zone, HPZ = Habitat Protection Zone, MUZ = Multiple Use Zone, SPZ = Special Purpose Zone."

# LaTeX will not break a line at an underscore, so a long campaign ID would run
# past its column. Swapping underscores for spaces lets the name wrap.
sampling_summary <- readRDS(here("data", park, "tidy",
                                 paste0(name, "_sampling-summary.RDS"))) %>%
  mutate(campaignid = gsub("_", " ", campaignid))

col_labels <- c("Date", "Campaign name", "Method", "Areas", "Samples",
                "Count", "Count", "Length")

sampling_summary %>%
  gt() %>%
  cols_label(dates         = col_labels[1],
             campaignid    = col_labels[2],
             method        = col_labels[3],
             areas         = col_labels[4],
             n_samples     = col_labels[5],
             habitat_count = col_labels[6],
             fish_count    = col_labels[7],
             fish_length   = col_labels[8]) %>%
  tab_spanner(label = "Habitat", columns = habitat_count) %>%
  tab_spanner(label = "Fish",    columns = c(fish_count, fish_length)) %>%
  sub_zero(zero_text = "-") %>%
  cols_width(dates         ~ pct(14),
             campaignid    ~ pct(20),
             method        ~ pct(11),
             areas         ~ pct(15),
             n_samples     ~ pct(10),
             habitat_count ~ pct(10),
             fish_count    ~ pct(10),
             fish_length   ~ pct(10)) %>%
  cols_align(align = "left",
             columns = c(dates, campaignid, areas)) %>%
  cols_align(align = "center",
             columns = c(method, n_samples, habitat_count, fish_count,
                         fish_length)) %>%
  opt_row_striping() %>%
  tab_style(style = cell_text(weight = "bold"),
            locations = list(cells_column_labels(),
                             cells_column_spanners(spanners = everything()))) %>%
  tab_options(table.width               = pct(100),
              table.font.size           = px(11),
              data_row.padding          = px(4),
              latex.use_longtable       = TRUE,
              column_labels.border.top.style = "none",
              table.border.top.style    = "none",
              table_body.hlines.style   = "none")
```

{{< pagebreak >}}

## Benthic ecosystem extent
### Observations

```{r}
#| fig-width: 10
#| fig-height: 6
#| fig-cap: "Habitat classes observed in spatially balanced stereo-BRUV and drop-camera imagery visualised as spatial pie charts. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
img <- png::readPNG(here("plots/ningaloo/habitat/ningaloo_scatterpie_2019_2021_2022_2024.png"))
grid::grid.raster(img)
```

{{< pagebreak >}}

### Dominant habitat types

```{r}
#| fig-width: 8
#| fig-height: 8
#| fig-cap: "Predicted dominant habitat type from stereo-BRUV groundtruthing. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
p_cat <- readRDS(here("plots/ningaloo/habitat/ningaloo_predicted-habitat-categorical_2019_2021_2022_2024.rds"))
p_cat
```

{{< pagebreak >}}

### Habitat distribution and probability

```{r}
#| fig-width: 8
#| fig-height: 7
#| fig-cap: "Predicted habitat distribution probability and uncertainty from stereo-BRUV groundtruthing. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
p_dom <- readRDS(here("plots/ningaloo/habitat/ningaloo_predicted-dominant-habitat_2019_2021_2022_2024.rds"))
p_dom
```

{{< pagebreak >}}

### Funcional reef distribution and probability

```{r}
#| fig-width: 8
#| fig-height: 7
#| fig-cap: "Predicted reef distribution probability and uncertainty from stereo-BRUV groundtruthing. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
img_path <- here("plots/ningaloo/habitat/ningaloo_predicted-reef-and-se_2019-2021-2022-2024.png")
img <- png::readPNG(img_path)
grid::grid.raster(img)
```

{{< pagebreak >}}

## Benchmarks of benthic natural values

```{=latex}
% Continue the running figure number into the benchmark sections, then switch
% to <metric>.<plot> numbering. `metric` is seeded to the last benthic-extent
% figure number so the majors carry on sequentially (4, 5, 6, ...).
\setcounter{metric}{\value{figure}}
\renewcommand{\thefigure}{A1.\arabic{metric}.\arabic{figure}}
```

<!-- Only sand and sessile invertebrates were modelled in 05_model-data_benthos.R
- macroalgae, rock and seagrasses never had enough non-zero samples (>80%
zeros) to fit. There are no control-plot chunks in this section: they are only
produced when combine_benthos is FALSE (years modelled separately), and here
the benthos years are pooled. -->

### Sand

```{=latex}
\stepcounter{metric}
```



```{r}
#| fig-width: 8
#| fig-height: 5.5
#| fig-cap: "Sand: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. Ningaloo Marine Park (Commonwealth). Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
img <- png::readPNG(here("plots/ningaloo/habitat/ningaloo_predicted-individual-habitat_sand_2019_2021_2022_2024.png"))
grid::grid.raster(img)
```

{{< pagebreak >}}

### Sessile invertebrates

```{=latex}
\stepcounter{metric}
```



```{r}
#| fig-width: 8
#| fig-height: 5.5
#| fig-cap: "Sessile invertebrates: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. Ningaloo Marine Park (Commonwealth). Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
img <- png::readPNG(here("plots/ningaloo/habitat/ningaloo_predicted-individual-habitat_sessile-invertebrates_2019_2021_2022_2024.png"))
grid::grid.raster(img)
```

{{< pagebreak >}}

## Benchmarks of demersal fish natural values

<!-- Fish years are pooled into a single combined prediction (see
06_model-data_fish.R), so there is no per-year fish control plot (SST-vs-year
comparison) anywhere in this section - it compares each year's prediction
against an SST time series, which has nothing to compare once there is only
one pooled prediction. The SAC and top-species bar plots below are built from
raw per-sample survey data, not the pooled model, and are still faceted by the
real survey years (2019, 2021, 2022, 2024). -->

### Survey effort

```{=latex}
\stepcounter{metric}
\renewcommand{\thefigure}{A1.\arabic{metric}}
```

```{r}
#| fig-width: 8
#| fig-height: 8
#| fig-cap: "Spatial distribution of observed species richness and total abundance (MaxN) per stereo-BRUV drop, pooled across all deployments. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters."
p_bubble <- readRDS(here("plots/ningaloo/fish/ningaloo_bubbleplot_richness-abundance.rds"))
p_bubble
```

{{< pagebreak >}}

```{=latex}
\renewcommand{\thefigure}{A1.\arabic{metric}.\arabic{figure}}
```

### Species richness

```{=latex}
\stepcounter{metric}
```



```{r}
#| fig-width: 6
#| fig-height: 3
#| fig-cap: "Demersal fish - Species richness: species accumulation curve by year from stereo-BRUV groundtruthing. Ningaloo Marine Park (Commonwealth)."
p_bar <- readRDS(here("plots/ningaloo/fish/ningaloo_SAC-sample.rds"))
p_bar
```

:::{.content-hidden when-format="html"}
### Species richness {.unlisted .unnumbered}
:::

```{r}
#| fig-width: 8
#| fig-height: 4
#| fig-cap: "Demersal fish - Species richness: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
img <- png::readPNG(here("plots/ningaloo/fish/ningaloo_predicted-individual-fish-metric_whole-assemblage_2019_2021_2022_2024.png"))
grid::grid.raster(img)
```

{{< pagebreak >}}

### Total abundance

```{=latex}
\stepcounter{metric}
```

```{r}
#| fig-width: 8
#| fig-height: 3.5
#| fig-cap: "Demersal fish - Total abundance: top 10 most abundant species (pooled across deployments) from stereo-BRUV groundtruthing. Ningaloo Marine Park (Commonwealth)."
p_bar <- readRDS(here("plots/ningaloo/fish/ningaloo_top_maxn_bar_plot.rds"))
p_bar
```


```{r}
#| fig-width: 8
#| fig-height: 4
#| fig-cap: "Demersal fish - Total abundance: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
img <- png::readPNG(here("plots/ningaloo/fish/ningaloo_predicted-individual-fish-metric_total-abundance_2019_2021_2022_2024.png"))
grid::grid.raster(img)
```

{{< pagebreak >}}

### Large Reef Fish Index\*

```{=latex}
\stepcounter{metric}
```



```{r}
#| fig-width: 8
#| fig-height: 3.7
#| fig-cap: "Demersal fish - Large Reef Fish Index\\*: top 10 greatest biomass species (pooled across deployments) from stereo-BRUV groundtruthing. Ningaloo Marine Park (Commonwealth). \\*Large Reef Fish Index representing the biomass of all fish greater than 20 cm (B20) has been adapted here for stereo-BRUV data by removing sharks and rays."
p_bar <- readRDS(here("plots/ningaloo/fish/ningaloo_top_b20_bar_plot.rds"))
p_bar
```

```{r}
#| fig-width: 8
#| fig-height: 4
#| fig-cap: "Demersal fish - Large Reef Fish Index\\*: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
img <- png::readPNG(here("plots/ningaloo/fish/ningaloo_predicted-individual-fish-metric_large-reef-fish-index_2019_2021_2022_2024.png"))
grid::grid.raster(img)
```

{{< pagebreak >}}

### Reef fish thermal index

```{=latex}
\stepcounter{metric}
```



```{r}
#| fig-width: 8.5
#| fig-height: 4
#| fig-cap: "Demersal fish - Reef Fish Thermal Index: species thermal niche of top 10 most abundant species (pooled across deployments) from stereo-BRUV groundtruthing. Ningaloo Marine Park (Commonwealth)."
p_bar <- readRDS(here("plots/ningaloo/fish/ningaloo_top_maxn_cti_bar_plot.rds"))
p_bar
```


```{r}
#| fig-width: 8
#| fig-height: 4
#| fig-cap: "Demersal fish - Reef Fish Thermal Index: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
img <- png::readPNG(here("plots/ningaloo/fish/ningaloo_predicted-individual-fish-metric_cti_2019_2021_2022_2024.png"))
grid::grid.raster(img)
```

{{< pagebreak >}}

## Threatened species

```{=latex}
\stepcounter{metric}
```

```{r}
#| label: tbl-threatened-species
#| tbl-cap: "Threatened species recorded in stereo-BRUV imagery across all survey years, showing IUCN Red List category, the number of samples each species was recorded in, and the total abundance (MaxN) summed across those samples. *This species is also listed as 'Conservation dependent' on the EPBC Act List of Threatened Fauna. ^This species is also listed as 'Vulnerable' on the EPBC Act List of Threatened Fauna."

threatened_species <- readRDS(here("data", park, "tidy",
                                   paste0(name, "_threatened-species.RDS")))

threatened_species %>%
  dplyr::select(species_label, australian_common_name, family, listing,
                n_samples, total_maxn, years_seen) %>%
  gt() %>%
  cols_label(species_label           = "Species",
             australian_common_name = "Common name",
             family                 = "Family",
             listing                = "IUCN category",
             n_samples              = "Samples",
             total_maxn             = "Total MaxN",
             years_seen             = "Years recorded") %>%
  sub_missing(missing_text = "-") %>%
  cols_width(species_label           ~ pct(21),
             australian_common_name ~ pct(19),
             family                 ~ pct(16),
             listing                ~ pct(14),
             n_samples              ~ pct(9),
             total_maxn             ~ pct(10),
             years_seen             ~ pct(11)) %>%
  cols_align(align = "left",
             columns = c(species_label, australian_common_name, family)) %>%
  cols_align(align = "center",
             columns = c(listing, n_samples, total_maxn, years_seen)) %>%
  tab_style(style = cell_text(style = "italic"),
            locations = cells_body(columns = species_label)) %>%
  opt_row_striping() %>%
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_column_labels()) %>%
  tab_options(table.width               = pct(100),
              table.font.size           = px(11),
              data_row.padding          = px(4),
              latex.use_longtable       = TRUE,
              column_labels.border.top.style = "none",
              table.border.top.style    = "none",
              table_body.hlines.style   = "none")
```

{{< pagebreak >}}

### Threatened shark and ray species

```{r}
#| fig-width: 18
#| fig-height: 18
#| fig-cap: "Occurrence of threatened, IUCN Red List category, sharks and rays in stereo-BRUV imagery. Points are scaled by abundance (MaxN) and coloured by survey year; grey points show every stereo-BRUV deployment. Commonwealth and coastal marine park boundaries are shown. The red line delimits coastal and Commonwealth waters."
readRDS(here("plots/ningaloo/fish/ningaloo_threatened-species-occurrence_page1_2019-2021-2022-2024.rds"))
```

{{< pagebreak >}}

<!-- 09_threatened-species.R paginates this plot at 12 species/page (4 cols x
3 rows) and prints how many pages it made. This assumes 1 page for the
threatened list and 1 page for the non-threatened list - check the printed
page counts and add/remove chunks below to match if they differ. fig-width/
fig-height are just a starting point - adjust freely, the .rds is a live
ggplot object so it re-flows to whatever size you give it. -->

### Other shark and ray species

```{r}
#| fig-width: 18
#| fig-height: 18
#| fig-cap: "Occurrence of shark and ray species that are not threatened in stereo-BRUV imagery. Points are scaled by abundance (MaxN) and coloured by survey year; grey points show every stereo-BRUV deployment, so a panel with no coloured points is a genuine absence rather than an absence of sampling. Commonwealth and state marine park boundaries are shown. The red line delimits state and commonwealth waters."
readRDS(here("plots/ningaloo/fish/ningaloo_non-threatened-species-occurrence_page1_2019-2021-2022-2024.rds"))
```
