---
title: "Appendix A - Abrolhos Shallow Bank Marine Park Natural Values Report"
subtitle: "North-west Marine Parks Network"
filters:
  - figure-numbering.lua
format:
  pdf:
    output-file: "Project 4.21-Abrolhos-Shallow-Bank-Appendix A-q-Natural values.pdf"
    fig-pos: 'H'
    geometry: 
      - top=2cm
      - bottom=2.5cm
      - left=2cm
      - right=2cm
    include-in-header:
      text: |
        \usepackage{caption}
        \usepackage{float}
        \usepackage{chngcntr}
        \floatplacement{table}{H}
        \newcounter{metric}
        \counterwithin*{figure}{metric}
        \captionsetup{format=plain, indention=0pt}
  html:
    output-file: "Project 4.21-Abrolhos-Shallow-Bank-Appendix A-q-Natural values.html"
    fig-dpi: 96
toc: true
execute:
  echo: false
  warning: false
  message: false
post-render: 10_move-pdf.R
---

```{r}
#| include: false
library(here)
library(ggplot2)
library(knitr)
library(terra)
library(tidyterra)
library(cowplot)
library(png)
library(grid)
library(tidytext)
library(ggtext)
library(dplyr)
library(gt)

# Park folder and file prefix - must match 00_config.yml
park <- "abrolhos_shallow-bank"
name <- "Shallow-bank"

# Survey years, and the two label styles used in output filenames.
#   pooled_label - benthos outputs, which are POOLED across years
#   dash_label   - the reef figure and the fish spatial metric figures
years        <- c("2021", "2025")
pooled_label <- paste(years, collapse = "_")   # e.g. "2021_2025"
dash_label   <- paste(years, collapse = "-")   # e.g. "2021-2025"

# Helper paths
hab  <- function(f) here("plots", park, "habitat", paste0(name, f))
fish <- function(f) here("plots", park, "fish",    paste0(name, f))

# Helper to drop a PNG into the document
show_png <- function(path) {
  grid::grid.raster(png::readPNG(path))
}
```

# Natural Values

Understanding of natural values

{{< pagebreak >}}

## Sampling summary

```{r}
#| label: tbl-sampling-effort
#| tbl-cap: "Summary of monitoring campaigns in the study area, showing the survey dates, sampling method, management zones sampled, and the number of samples with each data type. NPZ = National Park Zone, HPZ = Habitat Protection Zone, MUZ = Multiple Use Zone, SPZ = Special Purpose Zone."

# LaTeX will not break a line at an underscore, so a long campaign ID would run
# past its column. Swapping underscores for spaces lets the name wrap.
sampling_summary <- readRDS(here("data", park, "tidy",
                                 paste0(name, "_sampling-summary.RDS"))) %>%
  mutate(campaignid = gsub("_", " ", campaignid))

col_labels <- c("Date", "Campaign name", "Method", "Areas", "Samples",
                "Count", "Count", "Length")

sampling_summary %>%
  gt() %>%
  cols_label(dates         = col_labels[1],
             campaignid    = col_labels[2],
             method        = col_labels[3],
             areas         = col_labels[4],
             n_samples     = col_labels[5],
             habitat_count = col_labels[6],
             fish_count    = col_labels[7],
             fish_length   = col_labels[8]) %>%
  tab_spanner(label = "Habitat", columns = habitat_count) %>%
  tab_spanner(label = "Fish",    columns = c(fish_count, fish_length)) %>%
  sub_zero(zero_text = "-") %>%
  cols_width(dates         ~ pct(14),
             campaignid    ~ pct(20),
             method        ~ pct(11),
             areas         ~ pct(15),
             n_samples     ~ pct(10),
             habitat_count ~ pct(10),
             fish_count    ~ pct(10),
             fish_length   ~ pct(10)) %>%
  cols_align(align = "left",
             columns = c(dates, campaignid, areas)) %>%
  cols_align(align = "center",
             columns = c(method, n_samples, habitat_count, fish_count,
                         fish_length)) %>%
  opt_row_striping() %>%
  tab_style(style = cell_text(weight = "bold"),
            locations = list(cells_column_labels(),
                             cells_column_spanners(spanners = everything()))) %>%
  tab_options(table.width               = pct(100),
              table.font.size           = px(11),
              data_row.padding          = px(4),
              latex.use_longtable       = TRUE,
              column_labels.border.top.style = "none",
              table.border.top.style    = "none",
              table_body.hlines.style   = "none")
```

{{< pagebreak >}}

## Benthic ecosystem extent
### Observations

```{r}
#| fig-width: 9
#| fig-height: 6
#| fig-cap: "Habitat classes observed in spatially balanced stereo-BRUV and drop-camera imagery visualised as spatial pie charts, pooled across the 2021 and 2025 surveys. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
show_png(hab(paste0("_scatterpie_", pooled_label, ".png")))
```

{{< pagebreak >}}

### Dominant habitat types

```{r}
#| fig-width: 8
#| fig-height: 5
#| fig-cap: "Predicted dominant habitat type from stereo-BRUV and drop-camera groundtruthing, pooled across the 2021 and 2025 surveys. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
p_dom <- readRDS(hab(paste0("_predicted-dominant-habitat_", pooled_label, ".rds")))
p_dom
```

{{< pagebreak >}}

### Habitat distribution and probability

```{r}
#| fig-width: 8
#| fig-height: 5
#| fig-cap: "Predicted habitat class distribution from stereo-BRUV and drop-camera groundtruthing, pooled across the 2021 and 2025 surveys. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
p_cat <- readRDS(hab(paste0("_predicted-habitat-categorical_", pooled_label, ".rds")))
p_cat
```

{{< pagebreak >}}

### Functional reef distribution and probability

```{r}
#| fig-width: 9
#| fig-height: 5
#| fig-cap: "Predicted reef distribution probability and uncertainty from stereo-BRUV and drop-camera groundtruthing, pooled across the 2021 and 2025 surveys. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
show_png(hab(paste0("_predicted-reef-and-se_", dash_label, ".png")))
```

{{< pagebreak >}}

## Benchmarks of benthic natural values
### Sand

```{r}
#| fig-width: 9
#| fig-height: 5
#| fig-cap: "Sand: spatial plots of distribution probability and uncertainty from stereo-BRUV and drop-camera groundtruthing, pooled across the 2021 and 2025 surveys. ABMP = Abrolhos Marine Park (Commonwealth). Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
show_png(hab(paste0("_predicted-individual-habitat_sand_", pooled_label, ".png")))
```

{{< pagebreak >}}

### Macroalgae

```{r}
#| fig-width: 9
#| fig-height: 5
#| fig-cap: "Macroalgae: spatial plots of distribution probability and uncertainty from stereo-BRUV and drop-camera groundtruthing, pooled across the 2021 and 2025 surveys. ABMP = Abrolhos Marine Park (Commonwealth). Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
show_png(hab(paste0("_predicted-individual-habitat_macroalgae_", pooled_label, ".png")))
```

{{< pagebreak >}}

### Rock

```{r}
#| fig-width: 9
#| fig-height: 5
#| fig-cap: "Rock: spatial plots of distribution probability and uncertainty from stereo-BRUV and drop-camera groundtruthing, pooled across the 2021 and 2025 surveys. ABMP = Abrolhos Marine Park (Commonwealth). Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
show_png(hab(paste0("_predicted-individual-habitat_rock_", pooled_label, ".png")))
```

{{< pagebreak >}}

### Sessile invertebrates

```{r}
#| fig-width: 9
#| fig-height: 5
#| fig-cap: "Sessile invertebrates: spatial plots of distribution probability and uncertainty from stereo-BRUV and drop-camera groundtruthing, pooled across the 2021 and 2025 surveys. ABMP = Abrolhos Marine Park (Commonwealth). Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
show_png(hab(paste0("_predicted-individual-habitat_sessile-invertebrates_", pooled_label, ".png")))
```

{{< pagebreak >}}

## Benchmarks of demersal fish natural values

### Survey effort

```{r}
#| fig-width: 8
#| fig-height: 8
#| fig-cap: "Spatial distribution of observed species richness and total abundance (MaxN) per stereo-BRUV deployment, shown separately for the 2021 and 2025 surveys. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters."
p_bubble <- readRDS(fish("_bubbleplot_richness-abundance.rds"))
p_bubble
```

{{< pagebreak >}}

```{=latex}
% Switch to <metric>.<plot> numbering for the fish metrics, seeding `metric`
% to the last flat figure number so the majors carry on sequentially.
\setcounter{metric}{\value{figure}}
\renewcommand{\thefigure}{\arabic{metric}.\arabic{figure}}
```

### Species richness

```{=latex}
\stepcounter{metric}
```

```{r}
#| fig-width: 6
#| fig-height: 3
#| fig-cap: "Demersal fish - Species richness: benchmark plots by zone and depth contour from stereo-BRUV groundtruthing. ABMP = Abrolhos Marine Park (Commonwealth)."
p_control <- readRDS(fish("_control-plot_species-richness-per-bruv.rds"))
p_control
```

```{r}
#| fig-width: 7
#| fig-height: 4
#| fig-cap: "Demersal fish - Species richness: species accumulation curve by year and zoning status from stereo-BRUV groundtruthing. ABMP = Abrolhos Marine Park (Commonwealth)."
p_sac <- readRDS(fish("_SAC-sample.rds"))
p_sac
```

{{< pagebreak >}}

:::{.content-hidden when-format="html"}
### Species richness {.unlisted .unnumbered}
:::

```{r}
#| fig-width: 8
#| fig-height: 4
#| fig-cap: "Demersal fish - Species richness: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing, shown separately for the 2021 and 2025 surveys. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
show_png(fish(paste0("_predicted-individual-fish-metric_whole-assemblage_", dash_label, ".png")))
```

{{< pagebreak >}}

### Total abundance

```{=latex}
\stepcounter{metric}
```

```{r}
#| fig-width: 6
#| fig-height: 3
#| fig-cap: "Demersal fish - Total abundance: benchmark plots by zone and depth contour from stereo-BRUV groundtruthing. ABMP = Abrolhos Marine Park (Commonwealth)."
p_control <- readRDS(fish("_control-plot_total-abundance-per-bruv.rds"))
p_control
```

```{r}
#| fig-width: 10
#| fig-height: 4
#| fig-cap: "Demersal fish - Total abundance: top 10 most abundant species in each survey year from stereo-BRUV groundtruthing. ABMP = Abrolhos Marine Park (Commonwealth). Species appearing in only one year's top 10 are shown in bold."
p_bar <- readRDS(fish("_top_maxn_bar_plot.rds"))
p_bar
```

{{< pagebreak >}}

:::{.content-hidden when-format="html"}
### Total abundance {.unlisted .unnumbered}
:::

```{r}
#| fig-width: 8
#| fig-height: 4
#| fig-cap: "Demersal fish - Total abundance: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing, shown separately for the 2021 and 2025 surveys. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
show_png(fish(paste0("_predicted-individual-fish-metric_total-abundance_", dash_label, ".png")))
```

{{< pagebreak >}}

### Large Reef Fish Index\*

```{=latex}
\stepcounter{metric}
```

```{r}
#| fig-width: 6
#| fig-height: 3
#| fig-cap: "Demersal fish - Large Reef Fish Index\\*: benchmark plots by zone and depth contour from stereo-BRUV groundtruthing. ABMP = Abrolhos Marine Park (Commonwealth). \\*Large Reef Fish Index representing the biomass of all fish greater than 20 cm (B20) has been adapted here for stereo-BRUV data by removing sharks and rays."
p_control <- readRDS(fish("_control-plot_large-reef-fish-index-biomass-g-per-bruv.rds"))
p_control
```

```{r}
#| fig-width: 10
#| fig-height: 4
#| fig-cap: "Demersal fish - Large Reef Fish Index\\*: top 10 greatest biomass species in each survey year from stereo-BRUV groundtruthing. ABMP = Abrolhos Marine Park (Commonwealth). The 2021 survey is shown pooled across zones because only five deployments were in fished waters."
p_bar <- readRDS(fish("_top_b20_bar_plot_mixed.rds"))
p_bar
```

{{< pagebreak >}}

:::{.content-hidden when-format="html"}
### Large Reef Fish Index\* {.unlisted .unnumbered}
:::

```{r}
#| fig-width: 8
#| fig-height: 4
#| fig-cap: "Demersal fish - Large Reef Fish Index\\*: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing, shown separately for the 2021 and 2025 surveys. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
show_png(fish(paste0("_predicted-individual-fish-metric_large-reef-fish-index_", dash_label, ".png")))
```

{{< pagebreak >}}

### Reef fish thermal index

```{=latex}
\stepcounter{metric}
```

```{r}
#| fig-width: 6
#| fig-height: 3
#| fig-cap: "Demersal fish - Reef Fish Thermal Index: benchmark plots by zone and depth contour, with annual sea surface temperature (black line, grey band = SD) overlaid, from stereo-BRUV groundtruthing. ABMP = Abrolhos Marine Park (Commonwealth)."
p_control <- readRDS(fish("_control-plot_community-thermal-index-c.rds"))
p_control
```

```{r}
#| fig-width: 10
#| fig-height: 4
#| fig-cap: "Demersal fish - Reef Fish Thermal Index: species thermal niche of the top 10 most abundant species in each survey year from stereo-BRUV groundtruthing. ABMP = Abrolhos Marine Park (Commonwealth)."
p_bar <- readRDS(fish("_top_maxn_cti_bar_plot.rds"))
p_bar
```

{{< pagebreak >}}

:::{.content-hidden when-format="html"}
### Reef fish thermal index {.unlisted .unnumbered}
:::

```{r}
#| fig-width: 8
#| fig-height: 4
#| fig-cap: "Demersal fish - Reef Fish Thermal Index: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing, shown separately for the 2021 and 2025 surveys. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
show_png(fish(paste0("_predicted-individual-fish-metric_cti_", dash_label, ".png")))
```
