---
title: "Appendix A - Abrolhos Marine Park (Big Bank) Natural Values Report"
subtitle: "South-west Marine Parks Network"
filters:
  - figure-numbering.lua
format:
  pdf:
    output-file: "Project 4.21-Abrolhos-Big-Bank-2-Appendix A-q-Natural values.pdf"
    fig-pos: 'H'
    geometry: 
      - top=2cm
      - bottom=2.5cm
      - left=2cm
      - right=2cm
    include-in-header:
      text: |
        \usepackage{caption}
        \usepackage{float}
        \usepackage{chngcntr}
        \floatplacement{table}{H}
        \newcounter{metric}
        \counterwithin*{figure}{metric}
        \captionsetup{format=plain, indention=0pt}
  html:
    output-file: "Project 4.21-Abrolhos-Big-Bank-2-Appendix A-q-Natural values.html"
    fig-dpi: 96
toc: true
execute:
  echo: false
  warning: false
  message: false
post-render: 10_move-pdf.R
---

```{r}
#| include: false
library(here)
library(ggplot2)
library(knitr)
library(terra)
library(tidyterra)
library(cowplot)
library(png)
library(grid)
library(tidytext)
library(ggtext)
library(dplyr)
library(gt)

# Park folder and file prefix - must match 00_config.yml.
park <- "abrolhos_big-bank"
name <- "Big-bank"

# Survey years. Big Bank is a single-year (2021) survey, so the year suffix on
# the spatial-metric PNGs is just "2021" - script 08 saves them with
# paste(years, collapse = "-").
years     <- 2021
yr_suffix <- paste(years, collapse = "-")

# Path helpers - keeps the chunks below free of hardcoded park/name strings
hab  <- function(f) here("plots", park, "habitat", paste0(name, "_", f))
fish <- function(f) here("plots", park, "fish",    paste0(name, "_", f))

# Draw a saved PNG at full chunk size
draw_png <- function(path) grid::grid.raster(png::readPNG(path))
```

# Natural Values

Understanding of natural values

{{< pagebreak >}}

## Sampling summary

```{r}
#| label: tbl-sampling-effort
#| tbl-cap: "Summary of monitoring campaigns in the study area, showing the survey dates, sampling method, management zones sampled, and the number of samples with each data type. NPZ = National Park Zone, HPZ = Habitat Protection Zone, MUZ = Multiple Use Zone, SPZ = Special Purpose Zone."

# LaTeX will not break a line at an underscore, so a long campaign ID would run
# past its column. Swapping underscores for spaces lets the name wrap.
sampling_summary <- readRDS(here("data", park, "tidy",
                                 paste0(name, "_sampling-summary.RDS"))) %>%
  mutate(campaignid = gsub("_", " ", campaignid))

col_labels <- c("Date", "Campaign name", "Method", "Areas", "Samples",
                "Count", "Count", "Length")

sampling_summary %>%
  gt() %>%
  cols_label(dates         = col_labels[1],
             campaignid    = col_labels[2],
             method        = col_labels[3],
             areas         = col_labels[4],
             n_samples     = col_labels[5],
             habitat_count = col_labels[6],
             fish_count    = col_labels[7],
             fish_length   = col_labels[8]) %>%
  tab_spanner(label = "Habitat", columns = habitat_count) %>%
  tab_spanner(label = "Fish",    columns = c(fish_count, fish_length)) %>%
  sub_zero(zero_text = "-") %>%
  cols_width(dates         ~ pct(14),
             campaignid    ~ pct(20),
             method        ~ pct(11),
             areas         ~ pct(15),
             n_samples     ~ pct(10),
             habitat_count ~ pct(10),
             fish_count    ~ pct(10),
             fish_length   ~ pct(10)) %>%
  cols_align(align = "left",
             columns = c(dates, campaignid, areas)) %>%
  cols_align(align = "center",
             columns = c(method, n_samples, habitat_count, fish_count,
                         fish_length)) %>%
  opt_row_striping() %>%
  tab_style(style = cell_text(weight = "bold"),
            locations = list(cells_column_labels(),
                             cells_column_spanners(spanners = everything()))) %>%
  tab_options(table.width               = pct(100),
              table.font.size           = px(11),
              data_row.padding          = px(4),
              latex.use_longtable       = TRUE,
              column_labels.border.top.style = "none",
              table.border.top.style    = "none",
              table_body.hlines.style   = "none")
```

{{< pagebreak >}}

## Benthic ecosystem extent
### Observations

```{r}
#| fig-width: 8
#| fig-height: 4
#| fig-cap: "Habitat classes observed in spatially balanced stereo-BRUV and drop-camera imagery visualised as spatial pie charts. Australian Marine Park boundaries are shown. Bathymetric contours at 30, 70 and 200 m are shown."
draw_png(hab(paste0("scatterpie_", yr_suffix, ".png")))
```

{{< pagebreak >}}

### Dominant habitat types

```{r}
#| fig-width: 8
#| fig-height: 5
#| fig-cap: "Predicted dominant habitat type from stereo-BRUV groundtruthing. Australian Marine Park boundaries are shown. Bathymetric contours at 30, 70 and 200 m are shown."
p_cat <- readRDS(hab(paste0("predicted-habitat-categorical_", yr_suffix, ".rds")))
p_cat
```

{{< pagebreak >}}

### Habitat distribution and probability

```{r}
#| fig-width: 5.5
#| fig-height: 8
#| fig-cap: "Predicted habitat distribution probability and uncertainty from stereo-BRUV groundtruthing. Australian Marine Park boundaries are shown. Bathymetric contours at 30, 70 and 200 m are shown."
draw_png(hab(paste0("predicted-dominant-benthos-and-combined-se_", yr_suffix, ".png")))
```

{{< pagebreak >}}

### Functional reef distribution and probability

```{r}
#| fig-width: 10
#| fig-height: 4
#| fig-cap: "Predicted reef distribution probability and uncertainty from stereo-BRUV groundtruthing. Australian Marine Park boundaries are shown. Bathymetric contours at 30, 70 and 200 m are shown."
draw_png(hab(paste0("predicted-reef-and-se_", yr_suffix, ".png")))
```

{{< pagebreak >}}

## Benchmarks of benthic natural values

```{=latex}
% Continue the running figure number into the benchmark sections, then switch
% to <metric>.<plot> numbering. `metric` is seeded to the last benthic-extent
% figure number so the majors carry on sequentially (4, 5, 6, ...).
\setcounter{metric}{\value{figure}}
\renewcommand{\thefigure}{\arabic{metric}.\arabic{figure}}
```

### Sand

```{=latex}
\stepcounter{metric}
```

```{r}
#| fig-width: 7
#| fig-height: 3
#| fig-cap: "Sand: temporal plots by depth contour from stereo-BRUV groundtruthing. ABMP = Abrolhos Marine Park (Commonwealth)."
p_control <- readRDS(hab("control-plot_sand.rds"))
p_control
```

```{r}
#| fig-width: 10
#| fig-height: 5
#| fig-cap: "Sand: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. ABMP = Abrolhos Marine Park (Commonwealth). Australian Marine Park boundaries are shown. Bathymetric contours at 30, 70 and 200 m are shown."
draw_png(hab(paste0("predicted-individual-habitat_sand_", yr_suffix, ".png")))
```

{{< pagebreak >}}

### Sessile invertebrates

```{=latex}
\stepcounter{metric}
```

```{r}
#| fig-width: 7
#| fig-height: 3
#| fig-cap: "Sessile invertebrates: temporal plots by depth contour from stereo-BRUV groundtruthing. ABMP = Abrolhos Marine Park (Commonwealth)."
p_control <- readRDS(hab("control-plot_sessile-invertebrates.rds"))
p_control
```

```{r}
#| fig-width: 10
#| fig-height: 5
#| fig-cap: "Sessile invertebrates: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. ABMP = Abrolhos Marine Park (Commonwealth). Australian Marine Park boundaries are shown. Bathymetric contours at 30, 70 and 200 m are shown."
draw_png(hab(paste0("predicted-individual-habitat_sessile-invertebrates_", yr_suffix, ".png")))
```

{{< pagebreak >}}

## Benchmarks of demersal fish natural values

### Survey effort

```{=latex}
\stepcounter{metric}
\renewcommand{\thefigure}{\arabic{metric}}
```

```{r}
#| fig-width: 8
#| fig-height: 8
#| fig-cap: "Spatial distribution of observed species richness and total abundance (MaxN) per stereo-BRUV drop, pooled across all deployments. Australian Marine Park boundaries are shown."
p_bubble <- readRDS(fish("bubbleplot_richness-abundance.rds"))
p_bubble
```

{{< pagebreak >}}

```{=latex}
\renewcommand{\thefigure}{\arabic{metric}.\arabic{figure}}
```

### Species richness

```{=latex}
\stepcounter{metric}
```

```{r}
#| fig-width: 6
#| fig-height: 3
#| fig-cap: "Demersal fish - Species richness: benchmark plots by zone and depth contour from stereo-BRUV groundtruthing. ABMP = Abrolhos Marine Park (Commonwealth)."
p_control <- readRDS(fish("control-plot_species-richness-per-bruv.rds"))
p_control
```

```{r}
#| fig-width: 6
#| fig-height: 3
#| fig-cap: "Demersal fish - Species richness: species accumulation curve from stereo-BRUV groundtruthing. ABMP = Abrolhos Marine Park (Commonwealth)."
p_bar <- readRDS(fish("SAC-sample.rds"))
p_bar
```

{{< pagebreak >}}

:::{.content-hidden when-format="html"}
### Species richness {.unlisted .unnumbered}
:::

```{r}
#| fig-width: 6
#| fig-height: 7
#| fig-cap: "Demersal fish - Species richness: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. Australian Marine Park boundaries are shown. Bathymetric contours at 30, 70 and 200 m are shown."
draw_png(fish(paste0("predicted-individual-fish-metric_whole-assemblage_", yr_suffix, ".png")))
```

{{< pagebreak >}}

### Total abundance

```{=latex}
\stepcounter{metric}
```

```{r}
#| fig-width: 6
#| fig-height: 3
#| fig-cap: "Demersal fish - Total abundance: benchmark plots by zone and depth contour from stereo-BRUV groundtruthing. ABMP = Abrolhos Marine Park (Commonwealth)."
p_control <- readRDS(fish("control-plot_total-abundance-per-bruv.rds"))
p_control
```

```{r}
#| fig-width: 6
#| fig-height: 3
#| fig-cap: "Demersal fish - Total abundance: top 10 most abundant species (pooled across deployments) from stereo-BRUV groundtruthing. ABMP = Abrolhos Marine Park (Commonwealth)."
p_bar <- readRDS(fish("top_maxn_bar_plot.rds"))
p_bar
```

{{< pagebreak >}}

:::{.content-hidden when-format="html"}
### Total abundance {.unlisted .unnumbered}
:::

```{r}
#| fig-width: 6
#| fig-height: 7
#| fig-cap: "Demersal fish - Total abundance: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. Australian Marine Park boundaries are shown. Bathymetric contours at 30, 70 and 200 m are shown."
draw_png(fish(paste0("predicted-individual-fish-metric_total-abundance_", yr_suffix, ".png")))
```

{{< pagebreak >}}

### Large Reef Fish Index\*

```{=latex}
\stepcounter{metric}
```

```{r}
#| fig-width: 6
#| fig-height: 3
#| fig-cap: "Demersal fish - Large Reef Fish Index\\*: benchmark plots by zone and depth contour from stereo-BRUV groundtruthing. ABMP = Abrolhos Marine Park (Commonwealth). \\*Large Reef Fish Index representing the biomass of all fish greater than 20 cm (B20) has been adapted here for stereo-BRUV data by removing sharks and rays."
p_control <- readRDS(fish("control-plot_large-reef-fish-index-biomass-g-per-bruv.rds"))
p_control
```

```{r}
#| fig-width: 6
#| fig-height: 3
#| fig-cap: "Demersal fish - Large Reef Fish Index\\*: top 10 greatest biomass species (pooled across deployments) from stereo-BRUV groundtruthing. ABMP = Abrolhos Marine Park (Commonwealth)."
draw_png(fish("top_b20_bar_plot.png"))
```

{{< pagebreak >}}

:::{.content-hidden when-format="html"}
### Large Reef Fish Index\* {.unlisted .unnumbered}
:::

```{r}
#| fig-width: 6
#| fig-height: 7
#| fig-cap: "Demersal fish - Large Reef Fish Index\\*: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. Australian Marine Park boundaries are shown. Bathymetric contours at 30, 70 and 200 m are shown."
draw_png(fish(paste0("predicted-individual-fish-metric_large-reef-fish-index_", yr_suffix, ".png")))
```

{{< pagebreak >}}

### Reef fish thermal index

```{=latex}
\stepcounter{metric}
```

```{r}
#| fig-width: 6
#| fig-height: 3
#| fig-cap: "Demersal fish - Reef Fish Thermal Index: benchmark plots by zone and depth contour, with monthly sea surface temperature (black line, grey band = SD) overlaid, from stereo-BRUV groundtruthing. ABMP = Abrolhos Marine Park (Commonwealth)."
p_control <- readRDS(fish("control-plot_community-thermal-index-c.rds"))
p_control
```

```{r}
#| fig-width: 6
#| fig-height: 3
#| fig-cap: "Demersal fish - Reef Fish Thermal Index: species thermal niche of top 10 most abundant species (pooled across deployments) from stereo-BRUV groundtruthing. ABMP = Abrolhos Marine Park (Commonwealth)."
p_bar <- readRDS(fish("top_maxn_cti_bar_plot.rds"))
p_bar
```

{{< pagebreak >}}

:::{.content-hidden when-format="html"}
### Reef fish thermal index {.unlisted .unnumbered}
:::

```{r}
#| fig-width: 6
#| fig-height: 7
#| fig-cap: "Demersal fish - Reef Fish Thermal Index: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. Australian Marine Park boundaries are shown. Bathymetric contours at 30, 70 and 200 m are shown."
draw_png(fish(paste0("predicted-individual-fish-metric_cti_", yr_suffix, ".png")))
```
