---
title: "Appendix A - Eastern Recherche Marine Park Natural Values Report"
subtitle: "South-west Marine Parks Network"
filters:
  - figure-numbering.lua
format:
  pdf:
    output-file: "Project 4.21-Eastern Recherche-2-Appendix A-q-Natural values.pdf"
    fig-pos: 'H'
    geometry: 
      - top=2cm
      - bottom=2.5cm
      - left=2cm
      - right=2cm
    include-in-header:
      text: |
        \usepackage{caption}
        \usepackage{float}
        \usepackage{chngcntr}
        \newcounter{metric}
        \counterwithin*{figure}{metric}
        \captionsetup{format=plain, indention=0pt}
  html:
    output-file: "Project 4.21-Eastern Recherche-2-Appendix A-q-Natural values.html"
    fig-dpi: 96
toc: true
execute:
  echo: false
  warning: false
  message: false
post-render: 10_move-pdf.R
---

```{r}
#| include: false
library(here)
library(ggplot2)
library(knitr)
library(terra)
library(tidyterra)
library(cowplot)
library(png)
library(grid)
library(tidytext)
library(ggtext)

# Paths are built from the same config the analysis scripts use, so the park
# and name strings never have to be typed out in this document.
config <- yaml::read_yaml(here("r/eastern-recherche_v3/02_create-report_appendix-A-natural-values/00_config.yml"))
park   <- config$park          # eastern-recherche_v3
name   <- config$name          # eastern-rechercheAMP_v3
years  <- unlist(config$years) # 2022, 2025

# Benthos is pooled (combine_benthos: TRUE), so habitat outputs carry a single
# underscore-joined label. Fish outputs keep years separate and are saved with
# a hyphen-joined suffix.
benthos_suffix <- paste(years, collapse = "_")   # 2022_2025
yr_suffix      <- paste(years, collapse = "-")   # 2022-2025

hab  <- function(f) here(paste0("plots/", park, "/habitat/", name, "_", f))
fish <- function(f) here(paste0("plots/", park, "/fish/",    name, "_", f))

show_png <- function(path) {
  grid::grid.raster(png::readPNG(path))
}
```

# Natural Values

Understanding of natural values

## Benthic ecosystem extent
### Observations

```{r}
#| fig-width: 8
#| fig-height: 7
#| fig-cap: "Habitat classes observed in spatially balanced stereo-BRUV and drop-camera (BOSS) imagery visualised as spatial pie charts, pooled across the 2022 and 2025 surveys. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
show_png(hab(paste0("scatterpie_", benthos_suffix, ".png")))
```

{{< pagebreak >}}

### Dominant habitat types

```{r}
#| fig-width: 8
#| fig-height: 7
#| fig-cap: "Predicted dominant habitat type from stereo-BRUV and drop-camera groundtruthing, pooled across survey years. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
p_cat <- readRDS(hab(paste0("predicted-habitat-categorical_", benthos_suffix, ".rds")))
p_cat
```

{{< pagebreak >}}

### Habitat distribution and probability

```{r}
#| fig-width: 8
#| fig-height: 7
#| fig-cap: "Predicted dominant benthic habitat from stereo-BRUV and drop-camera groundtruthing, pooled across survey years. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
p_dom <- readRDS(hab(paste0("predicted-dominant-habitat_", benthos_suffix, ".rds")))
p_dom
```

{{< pagebreak >}}

### Functional reef distribution and probability

```{r}
#| fig-width: 6
#| fig-height: 7
#| fig-cap: "Predicted reef distribution probability and uncertainty from stereo-BRUV and drop-camera groundtruthing, pooled across survey years. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
show_png(hab(paste0("predicted-reef-and-se_", yr_suffix, ".png")))
```

{{< pagebreak >}}

## Benchmarks of benthic natural values

```{=latex}
% Continue the running figure number into the benchmark sections, then switch
% to <metric>.<plot> numbering. `metric` is seeded to the last benthic-extent
% figure number so the majors carry on sequentially (4, 5, 6, ...).
\setcounter{metric}{\value{figure}}
\renewcommand{\thefigure}{\arabic{metric}.\arabic{figure}}
```

### Sand

```{=latex}
\stepcounter{metric}
```

```{r}
#| fig-width: 4
#| fig-height: 7
#| fig-cap: "Sand: spatial plots of distribution probability and uncertainty from stereo-BRUV and drop-camera groundtruthing, pooled across survey years. ERMP = Eastern Recherche Marine Park (Commonwealth). Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
show_png(hab(paste0("predicted-individual-habitat_sand_", benthos_suffix, ".png")))
```

{{< pagebreak >}}

### Macroalgae

```{=latex}
\stepcounter{metric}
```

```{r}
#| fig-width: 6
#| fig-height: 7
#| fig-cap: "Macroalgae: spatial plots of distribution probability and uncertainty from stereo-BRUV and drop-camera groundtruthing, pooled across survey years. ERMP = Eastern Recherche Marine Park (Commonwealth). Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
show_png(hab(paste0("predicted-individual-habitat_macroalgae_", benthos_suffix, ".png")))
```

{{< pagebreak >}}

### Sessile invertebrates

```{=latex}
\stepcounter{metric}
```

```{r}
#| fig-width: 6
#| fig-height: 7
#| fig-cap: "Sessile invertebrates: spatial plots of distribution probability and uncertainty from stereo-BRUV and drop-camera groundtruthing, pooled across survey years. ERMP = Eastern Recherche Marine Park (Commonwealth). Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
show_png(hab(paste0("predicted-individual-habitat_sessile-invertebrates_", benthos_suffix, ".png")))
```

{{< pagebreak >}}

## Benchmarks of demersal fish natural values

### Survey effort

```{=latex}
\stepcounter{metric}
\renewcommand{\thefigure}{\arabic{metric}}
```

```{r}
#| fig-width: 8
#| fig-height: 8
#| fig-cap: "Spatial distribution of observed species richness and total abundance (MaxN) per stereo-BRUV deployment. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters."
p_bubble <- readRDS(fish("bubbleplot_richness-abundance.rds"))
p_bubble
```

{{< pagebreak >}}

```{=latex}
\renewcommand{\thefigure}{\arabic{metric}.\arabic{figure}}
```

### Species richness

```{=latex}
\stepcounter{metric}
```

```{r}
#| fig-width: 6
#| fig-height: 3
#| fig-cap: "Demersal fish - Species richness: benchmark plots by zone and depth contour from stereo-BRUV groundtruthing. ERMP = Eastern Recherche Marine Park (Commonwealth)."
p_control <- readRDS(fish("control-plot_species-richness-per-bruv.rds"))
p_control
```

```{r}
#| fig-width: 6
#| fig-height: 3
#| fig-cap: "Demersal fish - Species richness: species accumulation curve by year from stereo-BRUV groundtruthing. ERMP = Eastern Recherche Marine Park (Commonwealth)."
p_bar <- readRDS(fish("SAC-sample.rds"))
p_bar
```

{{< pagebreak >}}

:::{.content-hidden when-format="html"}
### Species richness {.unlisted .unnumbered}
:::

```{r}
#| fig-width: 7
#| fig-height: 7
#| fig-cap: "Demersal fish - Species richness: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
show_png(fish(paste0("predicted-individual-fish-metric_whole-assemblage_", yr_suffix, ".png")))
```

{{< pagebreak >}}

### Total abundance

```{=latex}
\stepcounter{metric}
```

```{r}
#| fig-width: 6
#| fig-height: 3
#| fig-cap: "Demersal fish - Total abundance: benchmark plots by zone and depth contour from stereo-BRUV groundtruthing. ERMP = Eastern Recherche Marine Park (Commonwealth)."
p_control <- readRDS(fish("control-plot_total-abundance-per-bruv.rds"))
p_control
```

```{r}
#| fig-width: 6
#| fig-height: 3
#| fig-cap: "Demersal fish - Total abundance: top 10 most abundant species (pooled across deployments) from stereo-BRUV groundtruthing. ERMP = Eastern Recherche Marine Park (Commonwealth)."
p_bar <- readRDS(fish("top_maxn_bar_plot.rds"))
p_bar
```

{{< pagebreak >}}

:::{.content-hidden when-format="html"}
### Total abundance {.unlisted .unnumbered}
:::

```{r}
#| fig-width: 7
#| fig-height: 7
#| fig-cap: "Demersal fish - Total abundance: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
show_png(fish(paste0("predicted-individual-fish-metric_total-abundance_", yr_suffix, ".png")))
```

{{< pagebreak >}}

### Large Reef Fish Index\*

```{=latex}
\stepcounter{metric}
```

```{r}
#| fig-width: 7
#| fig-height: 3
#| fig-cap: "Demersal fish - Large Reef Fish Index\\*: benchmark plots by zone and depth contour from stereo-BRUV groundtruthing. ERMP = Eastern Recherche Marine Park (Commonwealth). \\*Large Reef Fish Index representing the biomass of all fish greater than 20 cm (B20) has been adapted here for stereo-BRUV data by removing sharks and rays."
p_control <- readRDS(fish("control-plot_large-reef-fish-index-biomass-g-per-bruv.rds"))
p_control
```

```{r}
#| fig-width: 7
#| fig-height: 3
#| fig-cap: "Demersal fish - Large Reef Fish Index\\*: top 10 greatest biomass species (pooled across deployments) from stereo-BRUV groundtruthing. ERMP = Eastern Recherche Marine Park (Commonwealth)."
p_bar <- readRDS(fish("top_b20_bar_plot.rds"))
p_bar
```

{{< pagebreak >}}

:::{.content-hidden when-format="html"}
### Large Reef Fish Index\* {.unlisted .unnumbered}
:::

```{r}
#| fig-width: 7
#| fig-height: 7
#| fig-cap: "Demersal fish - Large Reef Fish Index\\*: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
show_png(fish(paste0("predicted-individual-fish-metric_large-reef-fish-index_", yr_suffix, ".png")))
```

{{< pagebreak >}}

### Reef fish thermal index

```{=latex}
\stepcounter{metric}
```

```{r}
#| fig-width: 7
#| fig-height: 3
#| fig-cap: "Demersal fish - Reef Fish Thermal Index: benchmark plots by zone and depth contour, with monthly sea surface temperature (black line, grey band = SD) overlaid, from stereo-BRUV groundtruthing. ERMP = Eastern Recherche Marine Park (Commonwealth)."
p_control <- readRDS(fish("control-plot_community-thermal-index-c.rds"))
p_control
```

```{r}
#| fig-width: 9
#| fig-height: 3
#| fig-cap: "Demersal fish - Reef Fish Thermal Index: species thermal niche of top 10 most abundant species (pooled across deployments) from stereo-BRUV groundtruthing. ERMP = Eastern Recherche Marine Park (Commonwealth)."
p_bar <- readRDS(fish("top_maxn_cti_bar_plot.rds"))
p_bar
```

{{< pagebreak >}}

:::{.content-hidden when-format="html"}
### Reef fish thermal index {.unlisted .unnumbered}
:::

```{r}
#| fig-width: 7
#| fig-height: 7
#| fig-cap: "Demersal fish - Reef Fish Thermal Index: spatial plots of distribution probability and uncertainty from stereo-BRUV groundtruthing. Commonwealth marine park boundaries are shown. The red line delimits state and commonwealth waters. Bathymetric contour at 30 m is shown."
show_png(fish(paste0("predicted-individual-fish-metric_cti_", yr_suffix, ".png")))
```
